The role of advection versus diffusion in intraventricular transport

### 2025-03-06 paper/main.tex
Edits up to 4.9

\subsection{Computational model of solute transport after photoconversion}
We simulate transport of photoconverted proteins within the brain ventricles by
modeling transport of a solute with concentration $c(\xx, t)$ for $\xx\in\Omega$ and
time $t > 0$ by the advection-diffusion equation
\begin{equation}
    \pdifft{c} + \nabla\cdot\JJ = 0 \quad \text{in } \Omega,
    \label{eq:adv_diff_strong}
\end{equation}
where the total concentration flux $\JJ$ is assumed to consist of an advective and a diffusive flux:
\begin{equation*}
    \JJ = c\uu - D\nabla c .
\end{equation*}
Here, $\uu$ is the CSF velocity field, governed by the Stokes equations~\Cref{eq:stokes_eqs},
and $D$ is the molecular diffusion coefficient of the solute.
Initially, we set $c(\xx, 0)=0$ for all $\xx\in\Omega$.
The ventricular walls are assumed to be impermeable (admitting \emph{no flux}: $\JJ\cdot\nn=0$),
except at the anterior/posterior boundary $\Gp$.
At the anterior/posterior boundary, we permit the solute concentration to be advected into
and out of the domain in a manner that ensures conservation of mass
(see~\Cref{subsec:adv_diff_eq_weak_form} for details).

@@

\subsection{Estimation of diffusion coefficients via the Stokes-Einstein relation}
Motivated by an interest in molecules that are significant in neural development,
we study the transport of molecules resembling the Dendra2 Fluorescent Protein (Dendra2)
used in this study, the Starmaker+Green Fluorescent Protein (STM-GFP) reported in
Jeong~\emph{et al.}~\cite{Jeong2024TheZebrafish}, and exosomes.
To estimate the diffusion coefficients of these, when not
available in the literature, we use the Stokes--Einstein relation
\begin{equation}
    D = \frac{k_B T}{6\pi \mu R},
    \label{eq:D_stokes_einstein}
\end{equation}
which is an expression for the diffusion coefficient of a spherical
molecule suspended in a viscous fluid~\cite{Einstein1905UberTeilchen}.
In~\Cref{eq:D_stokes_einstein}, $k_B = \num{1.38e-23}$ J/K is the Boltzmann constant,
$T = 310$ K is the absolute temperature, $\mu$ is the dynamic viscosity of the
solvent (in this case the CSF), and $R$ is the radius of the molecule.
For the exosomes, we calculate $D$ using~\Cref{eq:D_stokes_einstein}
with $R=150$ nm, which represents exosomes at the larger end of the
scale~\cite{Moghassemi2024ExtracellularDecade}.
Hence this radius covers the slowest diffusing exosomes,
a choice made because smaller exosomes would have diffusion coefficients
similar to the other molecules we consider.
For STM-GFP, we estimated the diffusion coefficient by extrapolating
the experimentally observed diffusion coefficient of
GFP~\cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells},
assuming that if $D$ scales with $1/R$ according to~\Cref{eq:D_stokes_einstein},
$D$ would scale with the cube root of the molecule mass~\cite{Goodhill1997DiffusionGuidance}.
The resulting diffusion coefficients used in the simulations
are $D_1$ (exosomes), $D_2$ (STM-GFP) and $D_3$ (Dendra2)~(\Cref{tab:diff_coeff_table}).

@@

\begin{table}[h!]
    \centering
    \caption{Molecular mass and diffusion coefficients of simulated solutes.
    STM and GFP denote Starmaker Protein and Green Fluorescent Protein, respectively.
    (*) The value was extrapolated based on the value of $D$ for
    GFP reported in the literature~\cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells},
    assuming that $D$ scales with the cube root of the molecule mass.
    (\textdagger) The value was estimated using the Stokes-Einstein relation~\Cref{eq:D_stokes_einstein}.}
    \label{tab:diff_coeff_table}
    \begin{tabular}{l|ccc}
        \toprule
        Molecule & Mass [kDa] & $D \ [\mathrm{m^2/s}]$ & Reference \\ 
        \midrule 
        Dendra2 Fluorescent Protein & 25.6~\cite{Gurskaya2006EngineeringLight}
        & $\num{1.15e-10}$  & \cite{GuraSadovsky2017MeasurementExpansion}\\
        GFP (saline aqueous solution) & 26.9~\cite{UniProt2024GFPUniProtKB}
        & $\num{8.70e-11}$ & \cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells} \\
        STM-GFP & 66.2~\cite{UniProt2024StmUniProtKB} + 26.9~\cite{UniProt2024GFPUniProtKB}
        & $\num{5.75e-11}$ & * \\
        Exosomes (radius of 150 nm) & -- & $\num{2.17e-12}$ & \textdagger \\
        \bottomrule
    \end{tabular}
\end{table}

>>>>

\subsection{Computational model of solute transport after photoconversion}
We simulate transport of photoconverted proteins within the brain ventricles by
modeling transport of a solute with concentration $c(\xx, t)$ for $\xx\in\Omega$ and
time $t > 0$ by the advection-diffusion equation
\begin{equation}
    \pdifft{c} + \nabla\cdot\JJ = 0 \quad \text{in } \Omega,
    \label{eq:adv_diff_strong}
\end{equation}
where the total concentration flux $\JJ$ is assumed to consist of an advective and a diffusive flux:
\begin{equation*}
    \JJ = c\uu - D\nabla c .
\end{equation*}
Here, $\uu$ is the CSF velocity field, governed by the Stokes equations (\Cref{eq:stokes_eqs}),
and $D$ is the molecular diffusion coefficient of the solute.
Initially, we set $c(\xx, 0)=0$ for all $\xx\in\Omega$.
The ventricular walls are assumed to be impermeable (admitting \emph{no flux}: $\JJ\cdot\nn=0$),
except at the anterior/posterior boundary $\Gp$.
At the anterior/posterior boundary, we permit the solute concentration to be advected into
and out of the domain in a manner that ensures conservation of mass
(see~\Cref{subsec:adv_diff_eq_weak_form} for details).

@@

\subsection{Estimation of diffusion coefficients via the Stokes-Einstein relation}
Motivated by an interest in molecules that are significant in neural development,
we study the transport of molecules resembling (i) the Dendra2 Fluorescent Protein (Dendra2)
used in the photoconversion, (ii) the Starmaker+Green Fluorescent Protein (STM-GFP) reported by
Jeong~\emph{et al.}~\cite{Jeong2024TheZebrafish}, and (iii) exosomes.
To estimate the diffusion coefficients of these, when not
available in the literature, we use the Stokes--Einstein relation
\begin{equation}
    D = \frac{k_B T}{6\pi \mu R},
    \label{eq:D_stokes_einstein}
\end{equation}
which is an expression for the diffusion coefficient of a spherical
molecule suspended in a viscous fluid~\cite{Einstein1905UberTeilchen}.
In~\Cref{eq:D_stokes_einstein}, $k_B = \num{1.38e-23}$ J/K is the Boltzmann constant,
$T = 310$ K is the absolute temperature, $\mu$ is the dynamic viscosity of the
CSF, and $R$ is the radius of the molecule.
For the exosomes, we calculate $D$ using~\Cref{eq:D_stokes_einstein}
with the relatively large $R=150$ nm~\cite{Moghassemi2024ExtracellularDecade}.
%Hence this radius covers the slowest diffusing exosomes,
%a choice made because smaller exosomes would have diffusion coefficients
%similar to the other molecules we consider.
For STM-GFP, we estimate the diffusion coefficient by extrapolating
the experimentally observed diffusion coefficient of
GFP~\cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells},
assuming that if $D$ scales with $1/R$ according to~\Cref{eq:D_stokes_einstein},
$D$ would scale with the cube root of the molecule mass~\cite{Goodhill1997DiffusionGuidance}.
The resulting diffusion coefficients
are $D_1$ (exosomes), $D_2$ (STM-GFP) and $D_3$ (Dendra2) as given by \Cref{tab:diff_coeff_table}.

@@

\begin{table}[h!]
    \centering
    \caption{Molecular mass and diffusion coefficients of simulated solutes.
    STM and GFP denote Starmaker Protein and Green Fluorescent Protein, respectively.
    (*) The value was extrapolated based on the value of $D$ for
    GFP reported in the literature~\cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells},
    assuming that $D$ scales with the cube root of the molecule mass.
    (\textdagger) The value was estimated using the Stokes-Einstein relation~\Cref{eq:D_stokes_einstein}.}
    \label{tab:diff_coeff_table}
    \begin{tabular}{lc|ccc}
        \toprule
        Molecule & Label & Mass [kDa]  & $D \ [\mathrm{m^2/s}]$ & Reference \\ 
        \midrule 
        Exosomes (radius of 150 nm) & $D_1$ & -- & $\num{2.17e-12}$ & \textdagger \\
        GFP (saline aqueous solution) & -- & 26.9~\cite{UniProt2024GFPUniProtKB} & $\num{8.70e-11}$ & \cite{Swaminathan1997PhotobleachingDiffusion, Potma2001ReducedCells} \\
        STM-GFP & $D_2$ & 66.2~\cite{UniProt2024StmUniProtKB} + 26.9~\cite{UniProt2024GFPUniProtKB} & $\num{5.75e-11}$ & * \\
        Dendra2 Fluorescent Protein & $D_3$ & 25.6~\cite{Gurskaya2006EngineeringLight} & $\num{1.15e-10}$ & \cite{GuraSadovsky2017MeasurementExpansion}\\
        \bottomrule
    \end{tabular}
\end{table}
